---
title: "Introduction to gpxtoolbox"
author: "Martin Chan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to gpxtoolbox}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

`gpxtoolbox` is an R package designed for processing, analysing, and visualising GPS track data from GPX files. It provides tools for calculating route metrics such as distance, elevation gain/loss, and speed, as well as visualisation capabilities for elevation profiles and route maps.

This vignette demonstrates how to use the key functions of the package.

# Installation

You can install the development version of `gpxtoolbox` from GitHub:

```r
# Install devtools if not already installed
install.packages("devtools")

# Install gpxtoolbox
devtools::install_github("martinchan/gpxtoolbox")
```

# Example Workflow

The package includes an example GPX file, `icc_intro_ride.gpx`, which is used in the following examples.

## Load the Package

```{r}
library(gpxtoolbox)
```

## Read a GPX File

Use the `read_gpx_track()` function to extract track points from a GPX file.

```{r}
# Path to the example GPX file
example_gpx_path <- system.file("extdata", "icc_intro_ride.gpx", package = "gpxtoolbox")

# Read the GPX file
track_data <- read_gpx_track(example_gpx_path)

# View the first few rows of the track data
head(track_data)
```

## Calculate Metrics

Use `calculate_distance()` and `calculate_elevation_stats()` to compute distances and elevation statistics.

```{r}
# Calculate distances
track_data <- calculate_distance(track_data)

# Calculate elevation statistics
track_data <- calculate_elevation_stats(track_data)

# View the updated track data
head(track_data)
```

## Generate Route Statistics

Use `calculate_route_stats()` to summarise route statistics such as total distance, elevation gain, and average speed.

```{r}
# Calculate route statistics
stats <- calculate_route_stats(track_data)

# Print the statistics
print(stats)
```

## Visualise the Route

Use `plot_route()` to create visualisations of the route, including an elevation profile and a route map.

```{r}
# Plot the route
plot_route(track_data)
```

## Analyse a GPX File in One Step

The `analyse_gpx()` function provides a high-level interface to process a GPX file and return statistics, plots, or processed data.

```{r}
# Analyse the example GPX file and get summary statistics
stats <- analyse_gpx(example_gpx_path, return = "stats")
print(stats)

# Generate a plot of the route
analyse_gpx(example_gpx_path, return = "plot")

# Get processed track points data
track_data <- analyse_gpx(example_gpx_path, return = "data")
```

# Working with Web Links

You can also analyse GPX files from web links (e.g., Strava or RideWithGPS). The function will automatically download the GPX file and process it.

```r
# Analyse a GPX file from a Strava link
# stats <- analyse_gpx("https://www.strava.com/routes/3193200014155699162")

# Analyse a GPX file from a RideWithGPS link
# stats <- analyse_gpx("https://ridewithgps.com/routes/39294504")
```

# Conclusion

`gpxtoolbox` provides a comprehensive set of tools for working with GPX files in R. Whether you're an outdoor enthusiast, athlete, or researcher, this package makes it easy to process, analyse, and visualise GPS track data.
